Approximate Bayesian Computation in Large Scale Structure: constraining the galaxy-halo connection

### 2017-03-07 ccppabc/text/revision/main.tex
clean up in tex directory

\documentclass[12pt, preprint]{aastex}
\usepackage[breaklinks,colorlinks, 
urlcolor=blue,citecolor=blue,linkcolor=blue]{hyperref}
\usepackage{graphicx}	% For figures
\usepackage{natbib}	% For citep and citep
\usepackage{amsmath}	% for \iint
\usepackage{mathtools}
\usepackage{bm}
\usepackage[breaklinks]{hyperref}	% for blackboard bold numbers
\usepackage{hyperref}
\usepackage{algorithmic,algorithm}
\hypersetup{colorlinks}
\usepackage{color}
\usepackage{morefloats}
\definecolor{darkred}{rgb}{0.5,0,0}
\definecolor{darkgreen}{rgb}{0,0.5,0}
\definecolor{darkblue}{rgb}{0,0,0.5}
\hypersetup{colorlinks, linkcolor=darkblue, filecolor=darkgreen, urlcolor=darkred, citecolor=darkblue}
\DeclareMathOperator*{\argmax}{arg\,max}

\newcommand{\todo}[1]{{\em \textcolor{red}{ #1}}}
\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}
\newcommand{\lang}{\langle}
\newcommand{\ra}{\rangle}
\newcommand{\vep}{\bm{\epsilon}}
\newcommand{\ep}{\epsilon}
\newcommand{\pars}{\vec{\theta}}
\newcommand{\dev}{\mathrm{d}}

\begin{document}
\title{Approximate Bayesian Computation in Large Scale Structure: constraining the galaxy-halo connection}
\author{ChangHoon~Hahn\altaffilmark{\dagger1}, Mohammadjavad~Vakili\altaffilmark{\dagger1}, Kilian~Walsh\altaffilmark{1}, Andrew~P.~Hearin\altaffilmark{2}, David~W.~Hogg\altaffilmark{1,3,4,5}, and~Duncan~Campbell\altaffilmark{6}}
\altaffiltext{$\dagger$}{These authors have contributed equally to the paper: \url{mjvakili@nyu.edu}, \url{chh327@nyu.edu}}
\altaffiltext{1}{Center for Cosmology and Particle Physics, 
                 Department of Physics, New York University, 4 Washington Pl., New York, NY}    
\altaffiltext{2}{Yale Center for Astronomy \& Astrophysics, Yale University, New Haven, CT}
\altaffiltext{3}{Simons Center for Data Analysis, 160 Fifth Avenue, New York, NY}
\altaffiltext{4}{Center for Data Science, New York University, 726 Broadway, New York, NY}
\altaffiltext{5}{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg, Germany}

\altaffiltext{6}{Department of Astronomy, Yale University, New Haven, CT}

\begin{abstract}
The standard approaches to Bayesian parameter inference in large scale structure (LSS) assume a Gaussian functional form (chi-squared form) for the likelihood. They are also typically restricted to measurements 
such as the two point correlation function. 
Likelihood free inferences such as Approximate Bayesian Computation (ABC) make inference possible without 
assuming any functional form for the likelihood, thereby relaxing the assumptions and restrictions of 
the standard approach. Instead it relies on a forward generative model of the data and a metric
for measuring the distance between the model and data. In this work, we 
demonstrate that ABC is feasible for LSS parameter inference by using it to constrain parameters of the 
halo occupation distribution (HOD) model for populating dark matter halos with galaxies.

@@

\section{Introduction}
Cosmology was revolutionized in the 1990s with the introduction of likelihoods---%
pro\-ba\-bil\-ities for the data given the theoretical model---%
for combining data from different surveys and performing principled inferences of
the cosmological parameters (\citealt{White:1996aa, Riess:1998aa}). 
Nowhere has this been more true than in cosmic microwave background (CMB) studies,
where it is nearly possible to analytically evaluate a likelihood function that
involves no (or minimal) approximations (\citealt{Oh:1999aa}, \citealt{Wandelt:2004aa},  
\citealt{Eriksen:2004aa}, \citealt{planckI, planckII}).

@@

\section{ABC at work}\label{sec:abcatwork}
With the methodology and the key components of ABC explained above, here we set out to 
demonstrate how ABC can be used to constrain HOD parameters. We start, in Section \ref{sec:mock_obv} 
by creating our ``observation''. We select a set of HOD parameters which we deem as the ``true'' 
parameters and run it through our forward model producing a catalog of galaxy positions 
which we treat as our observation. Then, in Section \ref{sec:abcpmc_spec}, we explain 
the distance metric and other specific choices we make for the ABC-PMC algorithm. 
Ultimately, we demonstrate the use of ABC in LSS, in Section \ref{sec:abc_results},
where we present the parameter constraints we get from our ABC analyses. 
Lastly, in order to both assess the quality of the ABC-PMC parameter inference and also 
discuss the assumptions of the standard Gaussian likelihood approach, we compare the 
ABC-PMC results to parameter constraints using the standard approach in Section 
\ref{sec:abcvsmcmc}.

%%%%%%%%%%%%%%%%%%%%% SUBSECTION %%%%%%%%%%%%%%%%%%%%%%%
% Mock Observations  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Mock Observations}\label{sec:mock_obv}
In generating our ``observations'', and more generally for our forward model, we adopt 
the HOD model from \cite{zheng07} where the expected number of galaxies populating a 
dark matter halo is governed by Eqs (\ref{eq:ncen}) and (\ref{eq:nsat}). For the parameters 
of the model used to generate the fiducial mock observations, we choose the \cite{zheng07} best-fit HOD parameters for the 
SDSS main galaxy sample with a luminosity threshold $M_{r} = -21$: 
\begin{center}
\renewcommand{\arraystretch}{1.5}
\begin{tabular}{ccccc} \hline \hline 
$\log M_{\rm min} \; \left[ h^{-1} M_{\odot} \right]$ & $\sigma_{\log M}$ & $\log M_{0} \; \left[ h^{-1} M_{\odot} \right]$ & $\log M_{1} \; \left[ h^{-1} M_{\odot} \right]$ & $\alpha$ \\ \hline
$12.79$ & $0.39$ & $11.92$ & $13.94$ & $1.15$ \\ \hline 
\end{tabular} \par
\end{center}
Since these parameters are used to generate the mock observation, they are the parameters
that we ultimately want to recover from our parameter inference. We refer to them as the
true HOD parameters. Plugging them into our forward model (Section \ref{sec:forwardmodel}), 
we generate a catalog of galaxy positions.

@@

For our ABC-PMC analysis using the observables $\ngalaxy$ and $\xigg$, our distance metric 
$\vec\rho = [\rho_n, \rho_\xi]$ while the distance metric for the ABC-PMC analysis using 
the observables $\ngalaxy$ and $\gmf$, is $\vec\rho = [\rho_n, \rho_\zeta]$.

To avoid any complications from the choice for our prior, we select uniform priors over all parameters aside from 
the scatter parameter $\sigma_{\log\;M}$, for which we choose a log-uniform prior. 
We list the range of our prior distributions in Table ~\ref{tab:prior}.

@@

%%%%%%%%%%%%%%%%%%TABLES%%%%%%%%%%%%
\begin{table*}
\begin{center}
  \label{tab:prior}
  \caption{{\bf Prior Specifications}: The prior probability distribution 
  and its range for each of the \cite{zheng07} HOD parameters. 
  All mass parameters are in unit of $h^{-1}M_\odot$}
\begin{tabular}{@{}lllll}
\\ \hline 
    HOD Parameter & & Prior & & Range \\ \hline
  $\alpha$ & & Uniform & & [0.8, 1.3] \\
  $\sigma_{\rm \log M}$ & & Log--Uniform & &  [0.1, 0.7] \\
   $\log M_{0}$   & & Uniform & &  [10.0, 13.0] \\
  $\log M_{min}$ & &   Uniform & &  [11.02, 13.02] \\
  $\log M_{1}$ & & Uniform & & [13.0, 14.0] \\
 \hline
  \end{tabular}
\end{center}
\end{table*}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Mock observation 2PCF and GMF
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{paper_data_observables.pdf}
\caption{The two-point correlation function $\xigg(r)$ (left) and group multiplicity 
function $\gmf(N)$ (right) summary statistics of the mock observations generated from 
the ``true'' HOD parameters described in Section \ref{sec:mock_obv}. The width of the 
shaded region corresponds to the square root of the covariance matrix diagonal 
elements (Eq. \ref{eq:cov}). In our ABC analysis, we treat the $\xigg(r)$ 
and $\gmf(N)$ above as the summary statistics of the observation.}
\label{fig:mock}
\end{center}
\end{figure*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ABC Pool Evolution plot 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\begin{center}
\includegraphics[width=0.95\textwidth]{paper_ABC_poolevolution_nbargmf.pdf}
\caption{We demonstrate the evolution of the ABC particles, $\pars_t$, over iterations $t = 1$ to $9$ 
in the $\log \mathcal{M}_{min}$ and $\log \mathcal{M}_1$ parameter space. 
$\bar{n}$ and $\gmf(N)$ are used as observables for the above results. For reference, 
in each panel, we include the ``true'' HOD parameters (black star) listed in 
Section \ref{sec:mock_obv}. The initial distance threshold, $\vec\epsilon_1 = 
[\infty, \infty]$ at $t=1$ (top left) so the $\pars_1$ spans the entire range of the prior distribution, 
which is also the range of the panels. We see for $t < 5$, the parameter space occupied 
by the ABC $\pars_t$ shrinks dramatically. Eventually when the algorithm converges, 
$t > 7$, the parameter space occupied by $\pars_t$ no longer shrinks and their 
distributions represent the posterior distribution of the 
parameters. At $t=9$, the final iteration, the ABC algorithm has converged and 
we find that $\pars_\mathrm{true}$ lies safely within the $68\%$ confidence region.
} 
\label{fig:pool_demo}
\end{center}
\end{figure*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ABC Pool Convergence 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{paper_ABCconvergence.pdf}
\caption{We illustrate the convergence of the ABC algorithm through the evolution 
of the ABC particle distribution as a function of iteration for 
parameters $\log \mathcal{M}_\mathrm{min}$ (left), $\alpha$ (center), and 
$\log \mathcal{M}_1$ (right).
The top panel corresponds our ABC results using the observables $(\bar{n}, \gmf(N))$, 
while the lower panel plots  corresponds to the ABC results using $(\bar{n}, \xigg(r))$. 
The distributions of parameters show no significant change after $t > 7$, which suggests
that the ABC algorithm has converged.}
\label{fig:abc_converge}
\end{center}
\end{figure*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ABC Corner Plots
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\begin{center}
\includegraphics[width=0.85\textwidth]{paper_ABCcorner_nbarxi.pdf}
\caption{We present the constraints on the \cite{zheng07} HOD model parameters
obtained from our ABC-PMC analysis using $\bar{n}$ and $\xigg(r)$ as observables.  
The diagonal panels plot the posterior distribution of each HOD parameter with 
vertical dashed lines marking the $50\%$ quantile and $68\%$ confidence intervals of the 
distribution. The off-diagonal panels plot the degeneracies between parameter pairs. 
The range of each panel corresponds to the range of our prior choice. The ``true''
HOD parameters, listed in Section \ref{sec:mock_obv}, are also plotted in each of 
the panels (black). For $\log\mathcal{M}_0$, $\alpha$, and $\sigma_{\log M}$, 
the ``true'' parameter values lie near the center of the $68\%$ confidence interval 
of the posterior distribution. For $\log\mathcal{M}_1$ and $\log\mathcal{M}_\mathrm{min}$, 
which have tight constraints, the ``true'' values lie within the $68\%$ confidence 
interval. Ultimately, the ABC parameter constraints we obtain in our analysis are consistent 
with the ``true'' HOD parameters. 
}
\label{fig:abc_corner_nbarxi}
\end{center}
\end{figure*}

\begin{figure*}
\begin{center}
\includegraphics[width=0.85\textwidth]{paper_ABCcorner_nbargmf.pdf}
\caption{Same as Figure \ref{fig:abc_corner_nbarxi} but for our ABC analysis 
using $\bar{n}$ and $\gmf(N)$ as observables. The ABC parameter constraints we obtain are
consistent with the ``true'' HOD parameters.}
\label{fig:abc_corner_nbargmf}
\end{center}
\end{figure*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ABC Posterior Observable  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{paper_ABCposterior.pdf}
\caption{We compare the ABC-PMC posterior prediction for the observables $\xigg(r)$ (left) and $\gmf(N)$ (right) (orange; Section \ref{sec:abc_results}) 
to $\xigg(r)$ and $\gmf(N)$ of the mock observation (black) in the top panels.
In the lower panels, we plot the ratio between the ABC-PMC posterior predictions for $\xigg$ and $\gmf$ 
to the mock observation $\xigg^\mathrm{obvs}$ and $\gmf^\mathrm{obvs}$. The darker 
and lighter shaded regions represent the $68\%$ and $95\%$ confidence regions of the posterior predictions, 
respectively. The error-bars represent the square root of the diagonal elements of the error 
covariance matrix (equation \ref{eq:cov}) of the mock observations. Overall, the observables 
drawn from the ABC-PMC posteriors are in good agreement with $\xigg$ and $\gmf$ of the 
mock observations. The lower panels demonstrate that for both observables, the error-bars 
of the mock observations lie within the $68\%$ confidence interval of the ABC-PMC posterior 
predictions.} 
\label{fig:abc_moneyplot}
\end{center}
\end{figure*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ABC vs MCMC Histograms 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\begin{center}
\includegraphics[width=1.\textwidth]{paper_ABCvsMCMC_nbarxi.pdf}
\caption{We compare the $\log \mathcal{M}_{\rm min}$, $\alpha$, and $\log \mathcal{M}_{1}$ parameter 
constraints from ABC-PMC (orange) to the constraints from the Gaussian Likelihood MCMC (blue)
using $\ngalaxy$ and $\xigg(r)$ as observables. The \emph{top} panels show the histograms of the 
marginalized posterior PDFs over the parameters. In the \emph{bottom} panels, we include box plots marking
the confidence intervals of the posterior distributions. The boxes represent the $68\%$ 
confidence interval while the ``whiskers'' represent the $95\%$ confidence interval. We also
plot the ``true'' HOD parameters with vertical black dashed line. Marginalized posterior PDFs 
obtained from the two methods are consistent with each other. The ABC-PMC constraints are slightly 
narrower for $\log \mathcal{M}_{\rm min}$, slightly wider for $\log \mathcal{M}_{1}$, and slightly 
less biased for $\alpha$.} 
\label{fig:hist_nbarxi}
\end{center}
\end{figure*}

>>>>

% mnras_template.tex 
%
% LaTeX template for creating an MNRAS paper
%
% v3.0 released 14 May 2015
% (version numbers match those of mnras.cls)
%
% Copyright (C) Royal Astronomical Society 2015
% Authors:
% Keith T. Smith (Royal Astronomical Society)

% Change log
%
% v3.0 May 2015
%    Renamed to match the new package name
%    Version number matches mnras.cls
%    A few minor tweaks to wording
% v1.0 September 2013
%    Beta testing only - never publicly released
%    First version: a simple (ish) template for creating an MNRAS paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Basic setup. Most papers should leave these options alone.
\documentclass[fleqn,usenatbib]{mnras}

% MNRAS is set in Times font. If you don't have this installed (most LaTeX
% installations will be fine) or prefer the old Computer Modern fonts, comment
% out the following line
\usepackage{newtxtext,newtxmath}
% Depending on your LaTeX fonts installation, you might get better results with one of these:
%\usepackage{mathptmx}
%\usepackage{txfonts}

% Use vector fonts, so it zooms properly in on-screen viewing software
% Don't change these lines unless you know what you are doing
\usepackage[T1]{fontenc}
\usepackage{ae,aecompl}

\usepackage{algorithmic,algorithm}

%%%%% AUTHORS - PLACE YOUR OWN PACKAGES HERE %%%%%

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%% AUTHORS - PLACE YOUR OWN COMMANDS HERE %%%%%

% Please keep new commands to a minimum, and use \newcommand not \def to avoid
% overwriting existing commands. Example:
%\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared

\newcommand{\todo}[1]{{\em \textcolor{red}{ #1}}}
\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}
\newcommand{\lang}{\langle}
\newcommand{\ra}{\rangle}
\newcommand{\vep}{\bm{\epsilon}}
\newcommand{\ep}{\epsilon}
\newcommand{\pars}{\vec{\theta}}
\newcommand{\dev}{\mathrm{d}}
\newcommand{\mstar}{h^{-1}M_\odot}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%% TITLE PAGE %%%%%%%%%%%%%%%%%%%

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[ABC in large scale structure]{Approximate Bayesian Computation in Large Scale Structure: constraining the galaxy-halo connection}

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[Hahn \& Vakili et al.]{
\parbox{\textwidth}{
ChangHoon~Hahn,$^{1,\ast}$\thanks{E-mail: chh327@nyu.edu}
Mohammadjavad~Vakili,$^{1,\ast}$\thanks{E-mail: mjvakili@nyu.edu}
Kilian~Walsh,$^{1}$\\
Andrew~P.~Hearin,$^{2}$
David~W.~Hogg,$^{1,3,4,5}$
and Duncan Campbell,$^{6}$
}
  \vspace*{10pt} \\
% List of institutions
$*$ These authors have contributed equally to the paper\\ 
$^{1}$Center for Cosmology and Particle Physics,Department of Physics, New York University, 726 Broadway, New York, NY\\
$^{2}$Yale Center for Astronomy \& Astrophysics, Yale University, New Haven, CT\\
$^{3}$Flatiron institute, 160 Fifth Avenue, New York, NY\\
$^{4}$Center for Data Science, New York University, 726 Broadway, New York, NY\\
$^{5}$Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg, Germany\\
$^{6}$Department of Astronomy, Yale University, New Haven, CT
}

% These dates will be filled out by the publisher
\date{Accepted XXX. Received YYY; in original form ZZZ}

% Enter the current year, for the copyright statements etc.
\pubyear{2015}

% Don't change these lines
\begin{document}
\label{firstpage}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\maketitle

% Abstract of the paper
\begin{abstract}
The standard approaches to Bayesian parameter inference in large scale structure (LSS) assume a Gaussian functional form (chi-squared form) for the likelihood. They are also typically restricted to measurements 
such as the two point correlation function. 
Likelihood free inferences such as Approximate Bayesian Computation (ABC) make inference possible without 
assuming any functional form for the likelihood, thereby relaxing the assumptions and restrictions of 
the standard approach. Instead it relies on a forward generative model of the data and a metric
for measuring the distance between the model and data. In this work, we 
demonstrate that ABC is feasible for LSS parameter inference by using it to constrain parameters of the 
halo occupation distribution (HOD) model for populating dark matter halos with galaxies.

@@

% Select between one and six entries from the list of approved keywords.
% Don't make up new ones.
\begin{keywords}
Large scale structure -- cosmology -- statistic
\end{keywords}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%% BODY OF PAPER %%%%%%%%%%%%%%%%%%

\section{Introduction}

Cosmology was revolutionized in the 1990s with the introduction of likelihoods---%
pro\-ba\-bil\-ities for the data given the theoretical model---%
for combining data from different surveys and performing principled inferences of
the cosmological parameters (\citealt{White:1996aa, Riess:1998aa}). 
Nowhere has this been more true than in cosmic microwave background (CMB) studies,
where it is nearly possible to analytically evaluate a likelihood function that
involves no (or minimal) approximations (\citealt{Oh:1999aa}, \citealt{Wandelt:2004aa},  
\citealt{Eriksen:2004aa}, \citealt{planckI, planckII}).

@@

\section{ABC at work}\label{sec:abcatwork}
With the methodology and the key components of ABC explained above, here we set out to 
demonstrate how ABC can be used to constrain HOD parameters. We start, in Section \ref{sec:mock_obv} 
by creating our ``observation''. We select a set of HOD parameters which we deem as the ``true'' 
parameters and run it through our forward model producing a catalog of galaxy positions 
which we treat as our observation. Then, in Section \ref{sec:abcpmc_spec}, we explain 
the distance metric and other specific choices we make for the ABC-PMC algorithm. 
Ultimately, we demonstrate the use of ABC in LSS, in Section \ref{sec:abc_results},
where we present the parameter constraints we get from our ABC analyses. 
Lastly, in order to both assess the quality of the ABC-PMC parameter inference and also 
discuss the assumptions of the standard Gaussian likelihood approach, we compare the 
ABC-PMC results to parameter constraints using the standard approach in Section 
\ref{sec:abcvsmcmc}.
%%%%%%%%%%%%%%%%%%%%% SUBSECTION %%%%%%%%%%%%%%%%%%%%%%%
% Mock Observations  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Mock Observations}\label{sec:mock_obv}
In generating our ``observations'', and more generally for our forward model, we adopt 
the HOD model from \cite{zheng07} where the expected number of galaxies populating a 
dark matter halo is governed by Eqs (\ref{eq:ncen}) and (\ref{eq:nsat}). For the parameters 
of the model used to generate the fiducial mock observations, we choose the \cite{zheng07} best-fit HOD parameters for the 
SDSS main galaxy sample with a luminosity threshold $M_{r} = -21$: 
\begin{center}
\renewcommand{\arraystretch}{1.5}
\begin{tabular}{ccccc} \hline \hline 
$\log M_{\rm min}$ & $\sigma_{\log M}$ & $\log M_{0}$ & $\log M_{1}$ & $\alpha$ \\ \hline
$12.79$ & $0.39$ & $11.92$ & $13.94$ & $1.15$ \\ \hline 
\end{tabular} \par
\end{center}
Since these parameters are used to generate the mock observation, they are the parameters
that we ultimately want to recover from our parameter inference. We refer to them as the
true HOD parameters. Plugging them into our forward model (Section \ref{sec:forwardmodel}), 
we generate a catalog of galaxy positions.

@@

For our ABC-PMC analysis using the observables $\ngalaxy$ and $\xigg$, our distance metric 
$\vec\rho = [\rho_n, \rho_\xi]$ while the distance metric for the ABC-PMC analysis using 
the observables $\ngalaxy$ and $\gmf$, is $\vec\rho = [\rho_n, \rho_\zeta]$.
To avoid any complications from the choice for our prior, we select uniform priors over all parameters aside from the scatter parameter $\sigma_{\log\;M}$, for which we choose a log-uniform prior. 
We list the range of our prior distributions in Table ~\ref{tab:prior}.

@@

%%%%%%%%%%%%%%%%%%TABLES%%%%%%%%%%%%

\begin{table}
	\centering
	\caption{{\bf Prior Specifications}: The prior probability distribution 
  and its range for each of the \citet{zheng07} HOD parameters. All mass parameters are in unit of $\mstar$}
	\label{tab:prior}
	\begin{tabular}{lcr} % four columns, alignment for each
		\hline
		HOD Parameter & Prior & Range\\
		\hline
		$\alpha$ & Uniform & [0.8, 1.3]\\
		$\sigma_{\rm \log M}$ & Log-Uniform & [0.1, 0.7]\\
		$\log M_{0}$ & Uniform & [10.0, 13.0]\\
        $\log M_{min}$ & Uniform & [11.02, 13.02]\\
        $\log M_{1}$ & Uniform & [13.0, 14.0]\\
		\hline
	\end{tabular}
\end{table}

@@

\section*{Acknowledgements}

We thank Jessie Cisewsky for reading and making valuable comments on the draft. We would also like to thank Michael R. Blanton, Jeremy R. Tinker, Uros Seljak, Layne Price, 
Boris Leidstadt, Alex Malz, Patrick McDonald, and Dan Foreman-Mackey for productive 
and insightful discussions. MV was supported by NSF grant AST-1517237. DWH was supported 
by NSF (grants IIS-1124794 and AST-1517237), NASA (grant NNX12AI50G), and the Moore-Sloan 
Data Science Environment at NYU. KW was supported by NSF grant AST-1211889. Computations 
were performed using computational resources at NYU-HPC. We thank Shenglong Wang, the 
administrator of NYU-HPC computational facility, for his consistent and continuous support 
throughout the development of this project. We would like to thank the organizers of 
the AstroHackWeek 2015 workshop (\url{http://astrohackweek.org/2015/}), 
since the direction and the scope of this investigation was ---to some degree--- initiated 
through discussions in this workshop. Throughout this investigation, we have made use of 
publicly available software packages $\mathtt{emcee}$ and $\mathtt{abcpmc}$. We have also used the publicly available python implementation of the FoF algorithm $\mathtt{pyfof}$ (\url{https://github.com/simongibbons/pyfof}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%% FIGURES %%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Mock observation 2PCF and GMF
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
%\begin{center}
\includegraphics[width=\textwidth]{paper_data_observables.pdf}
\caption{\label{fig:mock} The two-point correlation function $\xigg(r)$ (left) and group multiplicity 
function $\gmf(N)$ (right) summary statistics of the mock observations generated from 
the ``true'' HOD parameters described in Section \ref{sec:mock_obv}. The width of the 
shaded region corresponds to the square root of the covariance matrix diagonal 
elements (Eq. \ref{eq:cov}). In our ABC analysis, we treat the $\xigg(r)$ 
and $\gmf(N)$ above as the summary statistics of the observation.}
%\end{center}
\end{figure*}
%%%%%%%%%%%%%%%%%% End of FIGURE %%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ABC Pool Evolution plot 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
%\begin{center}
\includegraphics[width=0.95\textwidth]{paper_ABC_poolevolution_nbargmf.pdf}
\caption{\label{fig:pool_demo} We demonstrate the evolution of the ABC particles, $\pars_t$, over iterations $t = 1$ to $9$ in the $\log \mathcal{M}_{min}$ and $\log \mathcal{M}_1$ parameter space. 
$\bar{n}$ and $\gmf(N)$ are used as observables for the above results. For reference, 
in each panel, we include the ``true'' HOD parameters (black star) listed in 
Section \ref{sec:mock_obv}. The initial distance threshold, $\vec\epsilon_1 = 
[\infty, \infty]$ at $t=1$ (top left) so the $\pars_1$ spans the entire range of the prior distribution, 
which is also the range of the panels. We see for $t < 5$, the parameter space occupied 
by the ABC $\pars_t$ shrinks dramatically. Eventually when the algorithm converges, 
$t > 7$, the parameter space occupied by $\pars_t$ no longer shrinks and their 
distributions represent the posterior distribution of the 
parameters. At $t=9$, the final iteration, the ABC algorithm has converged and 
we find that $\pars_\mathrm{true}$ lies safely within the $68\%$ confidence region.} 
%\end{center}
\end{figure*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ABC Pool Convergence 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
%\begin{center}
\includegraphics[width=\textwidth]{paper_ABCconvergence.pdf}
\caption{\label{fig:abc_converge} We illustrate the convergence of the ABC algorithm through the evolution of the ABC particle distribution as a function of iteration for 
parameters $\log \mathcal{M}_\mathrm{min}$ (left), $\alpha$ (center), and 
$\log \mathcal{M}_1$ (right).
The top panel corresponds our ABC results using the observables $(\bar{n}, \gmf(N))$, 
while the lower panel plots  corresponds to the ABC results using $(\bar{n}, \xigg(r))$. 
The distributions of parameters show no significant change after $t > 7$, which suggests
that the ABC algorithm has converged.}
%\end{center}
\end{figure*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ABC Corner Plots
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
%\begin{center}
\includegraphics[width=0.85\textwidth]{paper_ABCcorner_nbarxi.pdf}
\caption{\label{fig:abc_corner_nbarxi} We present the constraints on the \citet{zheng07} HOD model parameters obtained from our ABC-PMC analysis using $\bar{n}$ and $\xigg(r)$ as observables.  
The diagonal panels plot the posterior distribution of each HOD parameter with 
vertical dashed lines marking the $50\%$ quantile and $68\%$ confidence intervals of the 
distribution. The off-diagonal panels plot the degeneracies between parameter pairs. 
The range of each panel corresponds to the range of our prior choice. The ``true''
HOD parameters, listed in Section \ref{sec:mock_obv}, are also plotted in each of 
the panels (black). For $\log\mathcal{M}_0$, $\alpha$, and $\sigma_{\log M}$, 
the ``true'' parameter values lie near the center of the $68\%$ confidence interval 
of the posterior distribution. For $\log\mathcal{M}_1$ and $\log\mathcal{M}_\mathrm{min}$, 
which have tight constraints, the ``true'' values lie within the $68\%$ confidence 
interval. Ultimately, the ABC parameter constraints we obtain in our analysis are consistent with the ``true'' HOD parameters.}
%\end{center}
\end{figure*}

\begin{figure*}
%\begin{center}
\includegraphics[width=0.85\textwidth]{paper_ABCcorner_nbargmf.pdf}
\caption{\label{fig:abc_corner_nbargmf}Same as Figure \ref{fig:abc_corner_nbarxi} but for our ABC analysis using $\bar{n}$ and $\gmf(N)$ as observables. The ABC parameter constraints we obtain are
consistent with the ``true'' HOD parameters.}
%\end{center}
\end{figure*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ABC Posterior Observable  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
%\begin{center}
\includegraphics[width=\textwidth]{paper_ABCposterior.pdf}
\caption{\label{fig:abc_moneyplot} We compare the ABC-PMC posterior prediction for the observables $\xigg(r)$ (left) and $\gmf(N)$ (right) (orange; Section \ref{sec:abc_results}) 
to $\xigg(r)$ and $\gmf(N)$ of the mock observation (black) in the top panels.
In the lower panels, we plot the ratio between the ABC-PMC posterior predictions for $\xigg$ and $\gmf$ 
to the mock observation $\xigg^\mathrm{obvs}$ and $\gmf^\mathrm{obvs}$. The darker 
and lighter shaded regions represent the $68\%$ and $95\%$ confidence regions of the posterior predictions, 
respectively. The error-bars represent the square root of the diagonal elements of the error 
covariance matrix (equation \ref{eq:cov}) of the mock observations. Overall, the observables 
drawn from the ABC-PMC posteriors are in good agreement with $\xigg$ and $\gmf$ of the 
mock observations. The lower panels demonstrate that for both observables, the error-bars 
of the mock observations lie within the $68\%$ confidence interval of the ABC-PMC posterior 
predictions.} 
%\end{center}
\end{figure*}
